---
title: "Monitoring Changes in Data"
author: "Mark van der Loo"
output:
  beamer_presentation:
    latex_engine: xelatex
    includes:
      in_header: ../00tex/presheader.tex
classoption: "aspectratio=169"
---

## Try the code

\centering\huge

`03input/monitoring.R`

## How to monitor changes

### It depends $\ldots$

- Cell-by-cell changes?
- Count changes?
- Count changes in satisfying rules?
- Measure changes in aggregates?

## Decomposing the number of changes in cells

\centering
\includegraphics[height=5cm]{fig/cellsplit.pdf}

```{r, eval=FALSE}
validate::cells(start=dataset1, step1=dataset2, step3=dataset3, ...)
```


## Decomposing the number of changes in validation results

\centering
\includegraphics[height=5cm]{fig/rulesplit.pdf}

```{r, eval=FALSE}
validate::compare(rules, start=dataset1, step1=dataset2, ...)
```


## Assignment

1. Read all the versions of datasets created today
2. Create plots showing the progress in the cell counts and rule counts methods.


# More on monitoring

## Process overview

\begin{center}
\includegraphics[]{fig/datastep.pdf}
\end{center}


## Process overview

\begin{center}
\includegraphics[]{fig/datastep2.pdf}
\end{center}


## How to measure changes between `data` and `data'`?

### Many ways

- List every change (record, variable, old, new) ('diff')
- Count differences in cells, or validation
- $\vdots$
- Note if something has changed (`TRUE`/`FALSE`)

## Needs

### Logging framework

- Supporting any type of comparison of `data` and `data'`
- Supporting any kind of transformation between `data` and `data'`
- Without demanding changes in the transforming functions
- That does not get in the way of the user

## Logging framework

### Idea

- A data cleaning procedure is implemented as a sequence of expressions (a script).
- These expressions are _composed_ into a programe when you run the script (`source()`)
- To obtain a logging framework that is not intrusive for the user, we can _change the way expressions are composed_.


## The `lumberjack` package

```{r, eval=FALSE}
# clean_it_up.R
library(validate)
dat <- read.csv("SBS2000.csv", stringsAsFactors = FALSE)

start_log(dat, logger=cellwise$new(key="id"))

dat <- errorlocate::replace_errors(dat,rules)
dat <- rspa::tag_missing(dat)
dat <- simputation::impute_rhd(dat, . ~ 1, backend="VIM")
dat <- rspa::match_restrictions(dat, rules)
write.csv(dat, file="output.csv", row.names=FALSE)
```

### In R, do

```{r,eval=FALSE}
library(lumberjack)
run_file("clean_it_up.R")
```



## Read the log:


\scriptsize{}
```{r}
read.csv("cellwise.csv") |> head(3)
```
\normalsize{}












